### 2016-11-23 local-capability-machine.tex
Changed the malloc specification so it leaves unused registers unchanged.

\begin{specification}[Malloc v.2]
  $c_\malloc$ satisfies the specification for malloc iff
  \begin{align*}
    &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
    &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \qquad \size = e'-b' \\
    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
  \end{align*}

@@

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\dominique{how is the return capability constructed: not on the stack, I presume, so this means mallocing a block of memory for the activation record?}
\lau{See the new addition below - when we don't have a stack, then we have no need to ``save'' local capabilities which means that we can have the return code as part of the program. The only thing we have to do is to have the capability point to the return address and make the capability into an entry capability. }
\lau{On further thought, conveniently the examples did not seem to need an activation record, but if we want \texttt{call} to work more generally speaking, then it needs to setup an activation record. I have changed the description of \texttt{call}, so it now uses an activation record. }
\dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
\lau{ I think this is a conceptual question as we can make it work with either. We already trust malloc to give us a fresh piece of memory and not reuse it later on, so we already assume malloc to be somewhat trusted, so why not go all the way? }
\dominique{what does ``fetch the capability ...'' mean?}
\lau{ We don't know where the capability resides, but if it is in memory, then it will be loaded into a register. }
\dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e. which do not need to be erased)?}
\dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}

>>>>

\begin{specification}[Malloc v.2]
  $c_\malloc$ satisfies the specification for malloc iff
  \begin{align*}
    &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
    &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
    &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
    &\qquad \qquad \qquad \size = e'-b' \land\\
  \end{align*}

@@

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\dominique{how is the return capability constructed: not on the stack, I presume, so  means mallocing a block of memory for the activation record?}
\lau{See the new addition below - when we don't have a stack, then we have no need to ``save'' local capabilities which means that we can have the return code as part of the program. The only thing we have to do is to have the capability point to the return address and make the capability into an entry capability. }
\lau{On further thought, conveniently the examples did not seem to need an activation record, but if we want \texttt{call} to work more generally speaking, then it needs to setup an activation record. I have changed the description of \texttt{call}, so it now uses an activation record. }
\dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
\lau{ I think this is a conceptual question as we can make it work with either. We already trust malloc to give us a fresh piece of memory and not reuse it later on, so we already assume malloc to be somewhat trusted, so why not go all the way? }
\dominique{what does ``fetch the capability ...'' mean?}
\lau{ We don't know where the capability resides, but if it is in memory, then it will be loaded into a register. }
\dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e. which do not need to be erased)?}
\dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}
